Data Migration under a Changing Schema

### 2023-08-12 2022Migration/articleMigrationFACS.tex
minor

\subsection{Data Migrations}
   Data migration occurs when a desired system replaces an existing one,
   while preserving the meaning of the present data as much as possible~\cite{Spivak2012}.
   %Just copying the set of data from the existing system to the desired system is obviously wrong if the schemas of both systems differ.
   %
   In practice, data migrations are done by deploying the desired system and the existing system side-by-side,
   while transferring data in a controlled fashion, as shown in Figure~\ref{fig:migration phase}.
\begin{figure}[bht]
   \begin{center}
     \includegraphics[scale=.35]{figures/datamigration-Migration-phase.png}
   \end{center}
\caption{Migration phase}
\label{fig:migration phase}
\end{figure}
   To automate the migration as much as possible,
   we also deploy a migration system,
   which has its own schema to specify the migration itself.
   The migration system comprises the schemas of both the existing system and the desired system,
   to implement the copying of data.
   It also implements the migration strategy,
   automated policies to resolve data pollution,
   and other rules at the discretion of the migration engineer.
   To fully automate every data migration is unrealistic.
   A migration engineer can minimize manual work by minimizing the size of the increments,
   but she cannot eliminate it completely.
   Data pollution, new business rules, or known issues in the existing system
   may occasionally require the migration engineer to tailor the migration script to specific needs.

   A data migration has the following distinct moments:
\begin{enumerate}
   \item the moment of transition (the MoT), i.e. the moment that users experience changed functionality;
   \item the moment of completion (the MoC), i.e. the migration engineer removes the migration system and the old system, including its data.
\end{enumerate}

   Before the migration, the existing system satisfies its invariants (by definition).
   When the migration system compiles correctly, it satisfies its own invariants,
   so it can be deployed in production.
   That is the MoT.
   The task of the migration system is to satisfy the invariants of the desired system.
   Some of these invariants can be satisfied automatically by means of enforce rules.
   There may be invariants, however, that require human interaction to resolve.
   Once all invariants of the desired system are satisfied, the migration is ready for the MoC.
   So, the schema of the migration system must contain the rules to satisfy all invariants of the desired system.

@@

   Let $\pair{\dataset}{\schema}$ be the existing system, and let ${\schema'}$ be the schema of the desired system.
   Then we define the migration systems $\migrsys$ and the events as follows:
\begin{enumerate}
\item We take a disjoint union of the relations, defining $\overleftarrow~$ and $\overrightarrow~$ as (overloaded) notation to generate new names, keeping the dataset of the old system and the rules of the new system:
   \begin{align}
   \overleftarrow{\declare{n}{x}{y}}&=\declare{(\id{old}.n)}{x}{y}&\overleftarrow{X} &= \{\overleftarrow{x}\mid x\in X\}\\
   \overrightarrow{\declare{n}{x}{y}}&=\declare{(\id{new}.n)}{x}{y}& \overrightarrow{X} &= \{\overrightarrow{x}\mid x\in X\}\\
   \triples_\migrsys &= \{\triple{a}{\overleftarrow{r}}{b} \mid \triple{a}{r}{b}\in\triples_\dataset\}&
   \instance_\migrsys &= \instance_\dataset\\
   \rels_\theenumi &= \overleftarrow\rels_{\schema} \cup \overrightarrow\rels_{\schema'}
   %&\rules_\theenumi &= \rules_\schema
   \end{align}
\item We create ENFORCE rules to copy equal relations, corresponding to equation~\ref{extraRelation} through~\ref{fill copies with new.isPartOf} in Section~\ref{sec:loading}:
   \begin{align}
   \widetilde{\declare{n}{x}{y}} &= \declare{(\id{copied}.n)}{x}{y}\\
   \rels_\theenumi &= \{\widetilde{r}\mid r \in \rels_\schema \cap \rels_{\schema'}\}\\
   \enforces_\theenumi &= \{\overrightarrow{r}\mapsfrom \popF{\overleftarrow{r} - \widetilde{r}} \mid \widetilde{r} \in \rels_2\} \cup \{\widetilde{r}\mapsfrom \popF{\overrightarrow{r}} \mid \widetilde{r} \in \rels_2\}
   \end{align}
\item For each rule, we generate two helper relations, {\tt viols} and {\tt fixed}. Note that Rule~\ref{warnings} only influences the user interface and therefore is not modeled here. This corresponds to equation~\ref{violRelation} through~\ref{rule uni}.
   \begin{align}
   \overleftarrow{\dataset'} &= \pair{\{\triple{x}{{r}}{y}\mid \triple{x}{\overleftarrow{r}}{y} \in \triples_{\dataset'}\}}{\instance_{\dataset'}}\\
   \overrightarrow{\dataset'} &= \pair{\{\triple{a}{{r}}{b} \mid \triple{a}{\overrightarrow{r}}{b}\in\triples_\dataset\}}{\instance_{\dataset'}} \\
   \rels_\theenumi &= \{{\tt viols}_u\mid u \in \rules_{\schema'}\} \cup \{{\tt fixed}_u\mid u \in \rules_{\schema'}\}\\
   \rules_\migrsys &= \{\lambda {\dataset'}.~ \viol{u}{\overrightarrow{\dataset'}} - \pop{{\tt viols}_u - {\tt fixed}_u}{\dataset'} \mid u\in\rules_{\schema'}\}\label{eqn:migrsysrules}\\
   \enforces_{\theenumi} &=\{{\tt viols}_u \mapsfrom \lambda \dataset'.~ \viol{u}{\overleftarrow{\dataset'}}\} \cup
    \{{\tt fixed}_u \mapsfrom \lambda \dataset'.~ {\tt viols}_u - \viol{u}{\overrightarrow{\dataset'}}\}\label{eqn:enforceForRules}
   \end{align}
\item Combining the above into a single migration system:
   \begin{align}
   \concepts_\migrsys &= \concepts_\dataset \cup  \concepts_{\dataset'}\\
   \rels_\migrsys &= \rels_1\cup\rels_2\cup \rels_3\\
   %\rules_\migrsys &= \rules_1\cup \rules_3\\
   \enforces_\migrsys &= \enforces_1\cup\enforces_2\cup\enforces_3
   \end{align}
\end{enumerate}

>>>>

\subsection{Data Migrations}
   Data migration occurs when a desired system replaces an existing one,
   while preserving the meaning of the present data as much as possible~\cite{Spivak2012}.
   %Just copying the set of data from the existing system to the desired system is obviously wrong if the schemas of both systems differ.
   %
   In practice, data migrations are done by deploying the desired system and the existing system side-by-side,
   while transferring data in a controlled fashion, as shown in Figure~\ref{fig:migration phase}.
\begin{figure}[bht]
   \begin{center}
     \includegraphics[scale=.35]{figures/datamigration-Migration-phase.png}
   \end{center}
\caption{Migration phase}
\label{fig:migration phase}
\end{figure}
   To automate the migration as much as possible,
   we also deploy a migration system,
   which has its own schema to specify the migration itself.
   The migration system comprises the schemas of both the existing system and the desired system,
   to implement the copying of data.
   To fully automate every data migration is unrealistic.
   A migration engineer can minimize manual work by minimizing the size of the increments,
   but she cannot eliminate it completely.
   Data pollution, new business rules, or known issues in the existing system
   may occasionally require the migration engineer to tailor the migration script to specific needs.

   A data migration has the following distinct moments\footnote{Completion can also be done gradually, but that is outside the scope of this paper.}:
\begin{enumerate}
   \item the moment of transition (the MoT), after which users experience changed functionality;
   \item the moment of completion (the MoC), after which the migration engineer removes the migration system and the old system, including its data.
\end{enumerate}

   Before the migration, the existing system satisfies its invariants (by definition).
   When the migration system compiles correctly, it satisfies its own invariants,
   so it can be deployed in production.
   That is the MoT.
   The task of the migration system is to satisfy the invariants of the desired system
   so that it can be deployed in production.
   For some of these invariants, enforcement rules can satisfy them automatically.
   There may be invariants, however, that require human interaction to resolve.
   Once all invariants of the desired system are satisfied, the migration is ready for the MoC.
   So, the schema of the migration system must contain the rules to satisfy all invariants of the desired system.

@@

   Let $\pair{\dataset}{\schema}$ be the existing system, and let ${\schema'}$ be the schema of the desired system.
   Then we define the migration systems $\migrsys$ and the events as follows:
\begin{enumerate}
\item We take a disjoint union of the relations, defining $\overleftarrow~$ and $\overrightarrow~$ as (overloaded) notation to generate new names, keeping the dataset of the old system and the rules of the new system:
   \begin{align}
   \overleftarrow{\declare{n}{x}{y}}&=\declare{(\id{old}.n)}{x}{y}&\overleftarrow{X} &= \{\overleftarrow{x}\mid x\in X\}\\
   \overrightarrow{\declare{n}{x}{y}}&=\declare{(\id{new}.n)}{x}{y}& \overrightarrow{X} &= \{\overrightarrow{x}\mid x\in X\}\\
   \triples_\migrsys &= \{\triple{a}{\overleftarrow{r}}{b} \mid \triple{a}{r}{b}\in\triples_\dataset\}&
   \instance_\migrsys &= \instance_\dataset\\
   \rels_\theenumi &= \overleftarrow\rels_{\schema} \cup \overrightarrow\rels_{\schema'}
   %&\rules_\theenumi &= \rules_\schema
   \end{align}
\item We create enforcement rules to copy equal relations, corresponding to equation~\ref{extraRelation} through~\ref{fill copies with new.isPartOf} in Section~\ref{sec:loading}:
   \begin{align}
   \widetilde{\declare{n}{x}{y}} &= \declare{(\id{copied}.n)}{x}{y}\\
   \rels_\theenumi &= \{\widetilde{r}\mid r \in \rels_\schema \cap \rels_{\schema'}\}\\
   \enforces_\theenumi &= \{\overrightarrow{r}\mapsfrom \popF{\overleftarrow{r} - \widetilde{r}} \mid \widetilde{r} \in \rels_2\} \cup \{\widetilde{r}\mapsfrom \popF{\overrightarrow{r}} \mid \widetilde{r} \in \rels_2\}
   \end{align}
\item For each rule, we generate two helper relations, {\tt viols} and {\tt fixed}. Note that Rule~\ref{warnings} only influences the user interface and therefore is not modeled here. This corresponds to equation~\ref{violRelation} through~\ref{rule uni}.
   \begin{align}
   \overleftarrow{\dataset'} &= \pair{\{\triple{x}{{r}}{y}\mid \triple{x}{\overleftarrow{r}}{y} \in \triples_{\dataset'}\}}{\instance_{\dataset'}}\\
   \overrightarrow{\dataset'} &= \pair{\{\triple{a}{{r}}{b} \mid \triple{a}{\overrightarrow{r}}{b}\in\triples_\dataset\}}{\instance_{\dataset'}} \\
   \rels_\theenumi &= \{{\tt viols}_u\mid u \in \rules_{\schema'}\} \cup \{{\tt fixed}_u\mid u \in \rules_{\schema'}\}\\
   \rules_\migrsys &= \{\lambda {\dataset'}.~ \viol{u}{\overrightarrow{\dataset'}} - \pop{{\tt viols}_u - {\tt fixed}_u}{\dataset'} \mid u\in\rules_{\schema'}\}\label{eqn:migrsysrules}\\
   \enforces_{\theenumi} &=\{{\tt viols}_u \mapsfrom \lambda \dataset'.~ \viol{u}{\overleftarrow{\dataset'}}\} \cup
    \{{\tt fixed}_u \mapsfrom \lambda \dataset'.~ {\tt viols}_u - \viol{u}{\overrightarrow{\dataset'}}\}\label{eqn:enforceForRules}
   \end{align}
\item Combining the above into a single migration system:
   \begin{align}
   \concepts_\migrsys &= \concepts_\dataset \cup  \concepts_{\dataset'}\\
   \rels_\migrsys &= \rels_1\cup\rels_2\cup \rels_3\\
   %\rules_\migrsys &= \rules_1\cup \rules_3\\
   \enforces_\migrsys &= \enforces_1\cup\enforces_2\cup\enforces_3
   \end{align}
\end{enumerate}
